---
title: "Homework  - Speeding on the I-5"
output:
  html_document:
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
show_solution = FALSE
```


A study investigated the distribution of passenger vehicle speeds traveling on the [Interstate 5 Freeway](https://en.wikipedia.org/wiki/Interstate_5) (I-5) in California. They discovered that the __distribution is nearly normal__ with a __mean of 72.6 miles/hour__ and a __standard deviation of 4.78 miles/hour__.




### Packages

We'll use the **tidyverse** package for data wrangling and visualisation.


You can load the package by running the following code:

```{r}
library(tidyverse) 
```


> Use R to compute all probabilities in the upcoming exercises. Beware aware that you submission NEEDS to contain ALL your R code as well as the corresponding output. So, NEVER choose the chunk option `echo=FALSE`.    



-------------


### Exercises 1

# The speed of passenger vehicles has a normal distribution with mean 72.6 and standard deviation 4.78. Hence, the probability is given by

```{r}
pnorm(80, mean = 72.6, sd = 4.78)
```


### Exercise 2

# We compute the probability of traveling less than 80 and then subtract the probability of traveling less than 60

```{r ex-2}
pnorm(80, mean = 72.6, sd = 4.78) - pnorm(60, mean = 72.6, sd = 4.78)
```

### Exercise 3

# The fastest 5% is nothing else than the 0.95 quantile of the distribution. We have two options to compute it. Either


```{r}
qnorm(0.95, mean = 72.6, sd = 4.78)
```

#### Exercise 4

#This is just the probability of traveling with a speed of more than 70 miles. We have again two options to compute the probability. Either we subtract from 1 the probability of traveling with a speed of less than 70 miles (subtract the cumulative distribution function at 70)


```{r}
1 - pnorm(70, mean = 72.6, sd = 4.78)
```




### Exercise 5  

# The officer is watching 5 Bernoulli trials (speeding yes/no) and counts the number of successes (car is speeding). Hence, the outcome follows a Binomial distribution with n=5 and success probability p=0.7067562. The question now is, how likely it is, that no car is speeding. 


```{r}
dbinom(0, size = 5, prob = 0.7067562)
```




### Exercise 6


# Solution: Now the officer is interested in the event of waiting until the first car is speeding. This event follows a geometric distribution with success probability p=0.7067562. From the lecture we now that expectation of this distribution is equal to 1/p, which leads to a value of
1 / 0.7067562



### Exercise 7

Start by reading in the data

```{r, eval=FALSE}
speed_data <- read_csv("speed_data.csv")
```


```{r ex-7-b, eval=FALSE}
speed_data |> 
  summarize(
    mean_speed = mean(speed),
    sd_speed = sd(speed)
  )
```
