---
title: "Smokers in Whickham"
author: "DD"
date: "Insert date here"
output: html_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
# you probably need to first install the mosaicData package
# do this in the console, not within this document 
library(mosaicData)  
```

### Exercise 1

#This is an observational study. The participants were not assigned to one of the conditions - smoker yes/no. It was merely observed if someone is  a smoker or not.

### Exercise 2

#The dataset contains 1314 observations. Since the data is given as a "pure" data.frame (and not a tibble), we find this information either in the help page or we run the code
```{r}
rnrow(Whickham)
```
# Each observations contains the infor,ation about the health outcome (dead/alive), the smoking status and age (at baseline) of one of the 1314 participants in the study.


### Exercise 3

# The dataset contains three variables: age, smoker, outcome. age is a numeric variable, smoker is a nominal variable and the type of outcome can be considered as ordinal (dead < alive) or just nominal.

# age: histogram
```{r}ggplot(Whickham, aes(x = age)) + 
  geom_histogram(colour = "white")
  # smoker: barplot
ggplot(Whickham, aes(x = smoker)) +
  geom_bar()
  # outcome: barplot
ggplot(Whickham, aes(x = outcome)) + 
  geom_bar()
````

### Exercise 4

# The expectation would be that among the deceased, the proportion of smokers would be higher than the proportion of non-smokers.

### Exercise 5


```{r}
Whickham |> 
  ggplot(aes(y = outcome, fill = smoker)) + 
  geom_bar()
```
# In the alive group we detect similar proportions for smokers and non-smokers, slightly higher for non-smokers. But in the dead group we observe a much larger proportion of dead non-smokers than dead smokers. This outcome is contradictory to our expectations.


```{r}
Whickham |> 
  select(smoker, outcome) |> 
  table() |> 
  prop.table(margin = 2) |> 
  addmargins()
```

### Exercise 6
We need to add a new variable to Whickham, so we will have to use mutate(). The value of the new variable must be determined using `case_when()`. A look at the help page of that function shows, that the given scheme is already of the correct form to be used in `case_when()`.
```{r}
Whickham <- Whickham |> 
  mutate(age_cat = case_when(
    age <= 44 ~ "18-44",
    age > 44 & age <= 64 ~ "45-64",
    age > 64 ~ "65+"
  ))
```


### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

Don't forget to label your code chunk. Do not use spaces in code chunk labels.
Don't forget to remove the chunk option `eval=FALSE` after you finished editing the code.

```{r, eval=FALSE}
Whickham |> 
  ggplot(aes(y = outcome, fill = smoker)) + 
  geom_bar() +
  facet_wrap(~ age_cat)
```

