---
title: "Homework - Smokers in Whickham"
author: "DD"
date: "Insert date here"
output: html_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
# you probably need to first install the mosaicData package
# do this in the console, not within this document 
library(mosaicData)  
```

### Exercise 1

Remove this text, and add your answer for Exercise 1 here.


### Exercise 2

Remove this text, and add your answer for Exercise 2 here.



### Exercise 3

Remove this text, and add your answer for Exercise 3 here.


Don't forget to remove the chunk option `eval=FALSE` after you finished editing the code.

```{r viz-variables, eval=FALSE}
ggplot(Whickham, aes(x = smoker)) +
  geom_ ___
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

Don't forget to label your code chunks. Do not use spaces in code chunk labels.
Don't forget to remove the chunk option `eval=FALSE` after you finished editing the code.

```{r, eval=FALSE}
Whickham |> 
  ggplot(aes(y = outcome, fill = ___)) + 
  ___
```


```{r, eval=FALSE}
Whickham |> 
  select(smoker, outcome) |> 
  table() |> 
  ___
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

Don't forget to label your code chunk. Do not use spaces in code chunk labels.
Don't forget to remove the chunk option `eval=FALSE` after you finished editing the code.

```{r, eval=FALSE}
Whickham <- Whickham |> 
  ___(age_cat = case_when(
    ___,
    ___,
    ___
  )
  )
```


### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

Don't forget to label your code chunk. Do not use spaces in code chunk labels.
Don't forget to remove the chunk option `eval=FALSE` after you finished editing the code.

```{r, eval=FALSE}
Whickham |> 
  count(smoker, age_cat, outcome) |> 
  group_by(outcome, ___) |>
  ___
```

